Studies of quantum dots

### 2016-11-12 Manuscript/paper.tex
Fix references

Several \textit{ab initio} methods have been applied to these systems, in
particular variational and diffusion Monte Carlo (VMC and DMC)
\cite{PhysRevB.68.035304,%
  PhysRevB.62.8120,PhysRevB.84.115302,PhysRevB.54.4780}, large-scale
diagonalization/full configuration interaction (FCI)
\cite{JJAP.36.3924,PhysRevB.56.6428,Kvaalcode,rontani:124102} and coupled
cluster (CC) theory \cite{PhysRevB.67.045320,heidari:114708,%
  PhysRevB.84.115302}.

@@

The harmonic oscillator (HO) is a ubiquitous system in quantum physics. It is the basis for an extremely diverse set of many-body calculations -- in nuclear physics \cite{HJORTHJENSEN1995125}, quantum chemistry \cite{Helgaker2002} [[missing ref for Helgaker 2002]], and quantum dot calculations of solid state physics \cite{reimann2002}. In these calculations, one typically seeks a few of the lowest eigenvalues $E_k$ and eigenvectors of a many-body Hamiltonian $\hat{H}$ that describes the full system by solving the Schr\"odinger equation,
\begin{equation*}
\hat{H} \ket{\psi_k} = E_k \ket{\psi_k}, \quad k \in \{1, 2, \ldots, k_{\text{max}}\}.
\end{equation*}
In all cases, the many-body wave function is expanded in a basis of
eigenstates of the one-body HO system.  Necessarily, in order to perform
numerical calculations the basis must be truncated to give an approximation.
In particular, the so-called ``curse of dimensionality'' implies that for a
system with many particles the number of available degrees of freedom per
particle is severely limited by the finiteness of a computer.

@@

The HO eigenfunctions are popular for several reasons. Many quantum systems, such as the quantum dot model considered here, can be considered perturbed harmonic oscillators \textit{per se}, thus the true eigenstates should be perturbations of the HO eigenstates.  Moreover, the HO has many beautiful properties, such as separability, invariance under orthogonal coordinate changes, and easily computable eigenfunctions, so that computing matrix elements of relevant operators becomes relatively simple.  Since HO eigenfunctions are defined on the \emph{entire} space in which particles live, a truncation of the spatial domain is unnecessary.  This avoids one of the main problems with approaches such as finite difference or finite element methods.\cite{RamMohan2002} [[Unknown citation]]

@@

The original choice suggested by Wegner \cite{PhysRepWegner0} reads
\begin{equation}
  \hat{\eta}^{\text{Wg}}
  = [\hat{H}^{\text{d}}, \hat{H} - \hat{H}^{\text{d}}]
  = [\hat{H}^{\text{d}}, \hat{H}]
  \label{eq:etaWegner}
\end{equation}
where $\hat{H}^{\text{d}}$ denotes the ``diagonal'' part of the Hamiltonian (and $\hat{H} - \hat{H}^{\text{d}}$ denotes the ``off-diagonal'' part) in the abstract sense described at the end of section \ref{subsubsec:srgmethods}.

@@

We have used IM-SRG to study the ground state energy of circular,
two-dimensional quantum dots up to $N=42$ particles. We utilized two
different generators, Wegner's and White's one, and realized that
Wegner's generator results in stiff equation systems, whereas White's
generator has much better numerics and is computationally more
efficient. Moreover, we found out that the use of an Hartree-Fock
basis improves numerical stability enormously, whereas analogous to
previous FCI \cite{Kvaalcode} and CCSD \cite{PhysRevB.84.115302}
calculations, the use of an effective interaction improves convergence
of the ground state energy as function of the number of oscillator
shells.  Our IM-SRG(2) results are in excellent agreement with other
many-body methods, in particular DMC, which can be considered as more
or less exact benchmark. For more than six particles, we lie in all
cases closer to the DMC result than corresponding CCSD calculations
do.

>>>>

Several \textit{ab initio} methods have been applied to these systems, in
particular variational and diffusion Monte Carlo (VMC and DMC)
\cite{PhysRevB.68.035304,%
  PhysRevB.62.8120,PhysRevB.84.115302,PhysRevB.54.4780}, large-scale
diagonalization/full configuration interaction (FCI)
\cite{JJAP.36.3924,PhysRevB.56.6428,2008arXiv0810.2644K,rontani:124102} and coupled
cluster (CC) theory \cite{PhysRevB.67.045320,heidari:114708,%
  PhysRevB.84.115302}.

@@

The harmonic oscillator (HO) is a ubiquitous system in quantum physics. It is the basis for an extremely diverse set of many-body calculations -- in nuclear physics \cite{HJORTHJENSEN1995125}, quantum chemistry \cite{helgaker2000molecular}, and quantum dot calculations of solid state physics \cite{reimann2002}. In these calculations, one typically seeks a few of the lowest eigenvalues $E_k$ and eigenvectors of a many-body Hamiltonian $\hat{H}$ that describes the full system by solving the Schr\"odinger equation,
\begin{equation*}
\hat{H} \ket{\psi_k} = E_k \ket{\psi_k}, \quad k \in \{1, 2, \ldots, k_{\text{max}}\}.
\end{equation*}
In all cases, the many-body wave function is expanded in a basis of
eigenstates of the one-body HO system.  Necessarily, in order to perform
numerical calculations the basis must be truncated to give an approximation.
In particular, the so-called ``curse of dimensionality'' implies that for a
system with many particles the number of available degrees of freedom per
particle is severely limited by the finiteness of a computer.

@@

The HO eigenfunctions are popular for several reasons. Many quantum systems, such as the quantum dot model considered here, can be considered perturbed harmonic oscillators \textit{per se}, thus the true eigenstates should be perturbations of the HO eigenstates.  Moreover, the HO has many beautiful properties, such as separability, invariance under orthogonal coordinate changes, and easily computable eigenfunctions, so that computing matrix elements of relevant operators becomes relatively simple.  Since HO eigenfunctions are defined on the \emph{entire} space in which particles live, a truncation of the spatial domain is unnecessary.  This avoids one of the main problems with approaches such as finite difference or finite element methods.\cite{ram2002finite} [[Unknown citation]]

@@

The original choice suggested by Wegner \cite{Wegner200177} reads
\begin{equation}
  \hat{\eta}^{\text{Wg}}
  = [\hat{H}^{\text{d}}, \hat{H} - \hat{H}^{\text{d}}]
  = [\hat{H}^{\text{d}}, \hat{H}]
  \label{eq:etaWegner}
\end{equation}
where $\hat{H}^{\text{d}}$ denotes the ``diagonal'' part of the Hamiltonian (and $\hat{H} - \hat{H}^{\text{d}}$ denotes the ``off-diagonal'' part) in the abstract sense described at the end of section \ref{subsubsec:srgmethods}.

@@

We have used IM-SRG to study the ground state energy of circular,
two-dimensional quantum dots up to $N=42$ particles. We utilized two
different generators, Wegner's and White's one, and realized that
Wegner's generator results in stiff equation systems, whereas White's
generator has much better numerics and is computationally more
efficient. Moreover, we found out that the use of an Hartree-Fock
basis improves numerical stability enormously, whereas analogous to
previous FCI \cite{2008arXiv0810.2644K} and CCSD \cite{PhysRevB.84.115302}
calculations, the use of an effective interaction improves convergence
of the ground state energy as function of the number of oscillator
shells.  Our IM-SRG(2) results are in excellent agreement with other
many-body methods, in particular DMC, which can be considered as more
or less exact benchmark. For more than six particles, we lie in all
cases closer to the DMC result than corresponding CCSD calculations
do.
